Toronto's Green Future: A Decade of Trends in City-Owned Renewable Energy Installations

### 2024-01-24 outputs/paper/paper.qmd
updated figure captions

library(tidyverse)
library(knitr)
library(ggplot2)
library(tmap)
library(sf)
```

@@

```{r}
#| label: tbl-cleaneddata
#| tbl-cap: "Cleaned Data of Renewable Energy Installations"
#| echo: false
#| message: false
#| warning: false

@@

```{r}
#| label: tbl-sizesummary
#| tbl-cap: "Size of Installation Summary Statistics"
#| echo: false

@@

```{r}
#| label: fig-sizehisto
#| fig-cap: "Distribution of size of installation"
#| echo: false
#| warning: false

cleaned_data %>% ggplot(aes (x= size)) + geom_histogram(bins=15) + labs(x="Size (kW)") + theme_minimal()
```
\newpage
### Type of Installation
Type of Installation (`type`) refers to whether the system falls under the FIT or microFIT program (with very few installations being classified as *storage*. FIT, or "Feed-In Tariff", is a program developed by the Ontario Power Authority which guarantees long-term financing and contracting for program participants, meant to incentivize the use of renewable energy sources [@FIT]. microFIT is part of the wider FIT program, but for smaller installations (10 kW or less).

@@

```{r}
#| label: fig-typebar
#| fig-cap: "Types of Installations"
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-yearinstallation
#| fig-cap: "Count of new installations per year"
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-locations
#| fig-cap: "Locations of renewable energy installations in Toronto"
#| echo: false
#| warning: false
toronto_map <- read_sf(here::here("inputs/data/map_data.gpkg"))

@@

```{r}
#| label: fig-yearsize
#| fig-cap: "Size of new installations over time with fitted regression line"
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-typeinstallation
#| fig-cap: "Count of new installations by type per year"
#| echo: false
#| warning: false

cleaned_data %>% ggplot(aes (x= factor(year), fill = type)) + geom_histogram(stat = "count") + labs(x="Year", fill="Type of Installation") 
```

\newpage
## Location of Installation over Time

@fig-yearlocation shows that even in earlier years, before the peak, installations were distributed fairly evenly across the city. Due to the low volume in these years, however, there were of course many gaps, which were strategically filled in later years. The map also reveals that there is a notable trend of expanding more into the city's periphery in recent years rather than installing more systems in the downtown core.

```{r}
#| label: fig-yearlocation
#| fig-cap: "Geographical Distribution of installations over the years"
#| echo: false
#| warning: false
tm_shape(toronto_map) +
  tm_borders() +  # Add borders for the Toronto map
  tm_shape(sf_data) +
  tm_dots(size = 1.5, col = "year", palette = "Blues", alpha = 0.7, title = "Year") +
  tm_layout(legend.position = c("right", "bottom"))

@@

>>>>

library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(tmap)
library(sf)
```

@@

```{r}
#| label: tbl-cleaneddata
#| tbl-cap: "Sample of Cleaned Data of City-Owned Renewable Energy Installations in Toronto"
#| echo: false
#| message: false
#| warning: false

@@

```{r}
#| label: tbl-sizesummary
#| tbl-cap: "Summary Statistics of the Size of City-Owned Renewable Energy Installations in Toronto in kilowatts (kW)"
#| echo: false

@@

```{r}
#| label: fig-sizehisto
#| fig-cap: "Distribution of the Size of City-Owned Renewable Energy Installations in Toronto in kilowatts (kW)"
#| fig-align: 'center'
#| echo: false
#| warning: false

cleaned_data %>% ggplot(aes (x= size)) + geom_histogram(bins=15) + labs(x="Size (kW)") + theme_minimal()
```
### Type of Installation
Type of Installation (`type`) refers to whether the system falls under the FIT or microFIT program (with very few installations being classified as *storage*. FIT, or "Feed-In Tariff", is a program developed by the Ontario Power Authority which guarantees long-term financing and contracting for program participants, meant to incentivize the use of renewable energy sources [@FIT]. microFIT is part of the wider FIT program, but for smaller installations (10 kW or less).

@@

```{r}
#| label: fig-typebar
#| fig-cap: "Count of Types of City-Owned Renewable Energy Installations in Toronto"
#| fig-align: 'center'
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-yearinstallation
#| fig-cap: "Count of new City-Owned Renewable Energy Installations in Toronto per year between 2010 and 2019"
#| fig-align: 'center'
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-locations
#| fig-cap: "Geographic distribution of City-Owned Renewable Energy Installations in Toronto"
#| fig-align: 'center'
#| echo: false
#| warning: false
toronto_map <- read_sf(here::here("inputs/data/map_data.gpkg"))

@@

```{r}
#| label: fig-yearsize
#| fig-cap: "Size of new City-Owned Renewable Energy Installations in Toronto between 2010 and 2019 with fitted regression line"
#| fig-align: 'center'
#| echo: false
#| warning: false

@@

```{r}
#| label: fig-typeinstallation
#| fig-cap: "Count of new City-Owned Renewable Energy Installations in Toronto by type of installation between 2010 and 2019"
#| fig-align: 'center'
#| echo: false
#| warning: false

cleaned_data %>% ggplot(aes (x= factor(year), fill = type)) + geom_histogram(stat = "count") + labs(x="Year", y="Count", fill="Type of Installation") 
```

\newpage
## Geographic Distribution of Installations

@fig-yearlocation shows that even in earlier years, before the peak, installations were distributed fairly evenly across the city. Due to the low volume in these years, however, there were of course many gaps, which were strategically filled in later years. The map also reveals that there is a notable trend of expanding more into the city's periphery in recent years rather than installing more systems in the downtown core. @fig-sizelocation (see @sec-appendix) additionally tells us that most larger installations are concentrated in the city's periphery.

```{r}
#| label: fig-yearlocation
#| fig-cap: "Geographical Distribution of City-Owned Renewable Energy Installations in Toronto between 2010 and 2019"
#| fig-align: 'center'
#| echo: false
#| warning: false

sf_data$year <- factor(sf_data$year)
tm_shape(toronto_map) +
  tm_borders() +  # Add borders for the Toronto map
  tm_shape(sf_data) +
  tm_dots(size= 2, col = "year" , palette = "Reds", alpha = 0.7, title = "Year") +
  tm_layout(legend.position = c("right", "bottom"))

@@

## Figures
```{r}
#| label: fig-sizelocation
#| fig-cap: "Geographical Distribution of City-Owned Renewable Energy Installations in Toronto by size"
#| echo: false
#| warning: false

sf_data <- sf_data %>% rename(Size = size)
tm_shape(toronto_map) +
  tm_borders() +  # Add borders for the Toronto map
  tm_shape(sf_data) +
  tm_dots(size= "Size" , col = "red", alpha = 0.7, title = "Size (kW)") +
  tm_layout(legend.position = c("right", "bottom"))
